cardinalR: Generating Interesting High-Dimensional Data Structures

### 2025-12-12 paper-cardinalR.Rmd
Fixed the figure caption

```{r, label = "fig-cluster-stats", fig.width=10, fig.height=3, out.width="100%", layout="l-body"}
# Examine the cluster stats
all_stats <- bind_rows(hc_stats, kmeans_stats, modelbased_stats)
all_stats |>
  pivot_longer(within.cluster.ss:sindex, names_to = "stat", values_to = "value") |>
  ggplot(aes(x=cl, y=value, colour=method)) +
    geom_line() +
    facet_wrap(~stat, ncol=3, scales="free_y") +
    xlab("Number of clusters") +
    ylab("") +
    theme_minimal() +
    theme(panel.border = element_rect(fill = 'transparent'),
          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
          axis.ticks.x = element_line(),
          axis.ticks.y = element_line(),
          legend.position = "bottom",
          axis.text.x = element_text(size = 7),
          axis.text.y = element_text(size = 7),
          axis.title.x = element_text(size = 7),
          axis.title.y = element_text(size = 7),
          plot.margin = margin(0, 0, 0, 0))
```

@fig-cluster-stats shows a selection of cluster metrics for $2-10$ clusters for each of the methods, $k$-means, hierarchical, model-based. As is typical, the suggestion of best solution varies between cluster statistics. There is some agreement that $4-5$ clusters is better than other choices (say why), and that $k$-means performs better than others (this might not hold when you add the next methods).

>>>>

```{r, label = "fig-cluster-stats", fig.width=10, fig.height=3, out.width="100%", layout="l-body", fig.cap="Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods."}

# Examine the cluster stats
all_stats <- bind_rows(hc_stats, kmeans_stats, modelbased_stats)
all_stats |>
  pivot_longer(within.cluster.ss:sindex, names_to = "stat", values_to = "value") |>
  ggplot(aes(x=cl, y=value, colour=method)) +
    geom_line() +
    facet_wrap(~stat, ncol=3, scales="free_y") +
    xlab("Number of clusters") +
    ylab("") +
    scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3')) +
    theme_minimal() +
    theme(panel.border = element_rect(fill = 'transparent'),
          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
          axis.ticks.x = element_line(),
          axis.ticks.y = element_line(),
          legend.position = "bottom",
          axis.text.x = element_text(size = 7),
          axis.text.y = element_text(size = 7),
          axis.title.x = element_text(size = 7),
          axis.title.y = element_text(size = 7),
          plot.margin = margin(0, 0, 0, 0))
```

Figure \@ref(fig:fig-cluster-stats) shows a selection of cluster metrics for $2-10$ clusters for each of the methods, $k$-means, hierarchical, model-based. As is typical, the suggestion of best solution varies between cluster statistics. There is some agreement that $4-5$ clusters is better than other choices (say why), and that $k$-means performs better than others.

### 2025-12-12 paper-cardinalR.tex
Fixed the figure caption

\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1}

\citet{fig-cluster-stats} shows a selection of cluster metrics for \(2-10\) clusters for each of the methods, \(k\)-means, hierarchical, model-based. As is typical, the suggestion of best solution varies between cluster statistics. There is some agreement that \(4-5\) clusters is better than other choices (say why), and that \(k\)-means performs better than others (this might not hold when you add the next methods).

>>>>

\begin{figure}[!]
\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1} \caption{Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.}\label{fig:fig-cluster-stats}
\end{figure}

Figure \ref{fig:fig-cluster-stats} shows a selection of cluster metrics for \(2-10\) clusters for each of the methods, \(k\)-means, hierarchical, model-based. As is typical, the suggestion of best solution varies between cluster statistics. There is some agreement that \(4-5\) clusters is better than other choices (say why), and that \(k\)-means performs better than others.
